\addcontentsline{toc}{chapter}{Acknowledgments}
\thispagestyle{empty}

\vspace*{20mm}

\begin{center}
    {\usekomafont{sectioning}\usekomafont{section} Acknowledgments}
\end{center}

\vspace{10mm}

First and foremost, I would like to express my deepest gratitude to my advisors, Takuto Asakura and Miriam Anschütz, for guiding me through this thesis's challenging yet rewarding journey. Your belief in me made all the difference.

Special thanks are also to Prof. Dr. Yusuke Miyao and Prof. Dr. Georg Groh for their supervision and for providing the facilities and environment I needed to focus on my work.

I extend my heartfelt thanks to Shoko Yamagishi, the secretary at Miyao Lab, for her incredible efficiency in navigating the bureaucracy of being an exchange student in Japan. You made my life easier in so many ways.

Words cannot adequately express my thanks to Aleksandra, who has been my pillar of strength, emotionally and physically. Your love and support have been unfaltering, even during the most stressful times.

My gratitude also goes out to Christoph, whose invaluable advice and insights made the daunting task during my entire studies seem manageable.

A warm thank you to Valentin, who was there when I needed him at the most critical moments. Your timely assistance made a world of difference.

I am also indebted to my family members Arun, Pallavi, and Parul, whose constant support and encouragement have been a steady force behind my efforts.

I would also like to extend my thanks to all the members of Panikecke\footnote{\url{https://discord.gg/RURaVwRe}} for their camaraderie and emotional support, as well as GPT and Rune S{\ae}tre (NTNU) for their meticulous proof-reading of my thesis.

Lastly, a big thank you to all my other friends Dhia, Philipp, Leonard, Daniel, and Mrinal, who have offered their support in many different ways. Every bit helped, and I am grateful for it.

\cleardoublepage{}
